---
title: "report"
date: "2024-12-02"
output:
  github_document:
    toc: true
    toc_float: true
    theme: sandstone
    code_folding: hide
---

```{r, warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(rvest)
library(dplyr)
library(tidyverse)
library(httr)
library(ggplot2)
library(tidyr)
library(maps)
library(knitr)
library(viridis)

theme_set(theme_minimal() + theme(legend.position = "right"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

**Project Name:** *Play by Numbers: Esports Trends Unveiled*

**Project Members:** Xun Sun (xs2569), Xinghao Qiao (xq2241), Huachen Shan (hs3478), Kanye Xu (kx2224), Fangchi Lu (fl2714) 

## Motivation

This project aims to analyze the evolving landscape of e-sports, focusing on patterns in player earnings and engagement strategies across demographics and game types. 

## Research Questions

## Data Wrangling

Access the top 100 players with Highest Overall Earnings

```{r}

# esports_earnings_top100_overall
url <- 'https://www.esportsearnings.com/players'
webpage <- read_html(url)

rank <- webpage %>% html_nodes("tr td:nth-child(1)") %>% html_text(trim = TRUE)
player_id <- webpage %>% html_nodes("tr td:nth-child(2)") %>% html_text(trim = TRUE)
player_name <- webpage %>% html_nodes("tr td:nth-child(3)") %>% html_text(trim = TRUE)
total_earnings <- webpage %>% html_nodes("tr td:nth-child(4)") %>% html_text(trim = TRUE)
highest_paying_game <- webpage %>% html_nodes("tr td:nth-child(5)") %>% html_text(trim = TRUE)
total_game_earnings <- webpage %>% html_nodes("tr td:nth-child(6)") %>% html_text(trim = TRUE)
percent_of_total <- webpage %>% html_nodes("tr td:nth-child(7)") %>% html_text(trim = TRUE)


data_overall <- data.frame(
  Rank = rank,
  Player_ID = player_id,
  Player_Name = player_name,
  Total_Earnings = total_earnings,
  Highest_Paying_Game = highest_paying_game,
  Total_Game_Earnings = total_game_earnings,
  Percent_of_Total = percent_of_total,
  stringsAsFactors = FALSE
)


data_overall$Rank <- as.numeric(gsub("\\.", "", data_overall$Rank))
head(data_overall)
write.csv(data_overall, "esports_earnings_top100_overall.csv", row.names = FALSE)

```

Access the top countries of players in esports who won prize money based on information published on the internet

```{r}

url <- 'https://www.esportsearnings.com/countries'
webpage <- read_html(url)

rows <- webpage %>% html_nodes("tr")

data <- data.frame(
  Rank = character(),
  Country = character(),
  Total_Earnings = character(),
  Number_of_Players = character(),
  Top_Game = character(),
  Earnings_From_Top_Game = character(),
  Percent_From_Top_Game = character(),
  stringsAsFactors = FALSE
)

for (row in rows) {
  
  columns <- row %>% html_nodes("td") %>% html_text(trim = TRUE)
  if (length(columns) == 7) {
    data <- bind_rows(data, as.data.frame(t(columns), stringsAsFactors = FALSE))
  }
}
colnames(data) <- c("Rank", "Country", "Total_Earnings", "Number_of_Players", 
                    "Top_Game", "Earnings_From_Top_Game", "Percent_From_Top_Game")

data[, 1:7] <- data[, 8:14]
colnames(data)[1:7] <- c("Rank", "Country", "Total_Earnings", "Number_of_Players", 
                         "Top_Game", "Earnings_From_Top_Game", "Percent_From_Top_Game")

data <- data[, 1:7]
data$Rank <- as.numeric(gsub("\\.", "", data$Rank))
data$Total_Earnings <- as.numeric(gsub("[^0-9\\.]", "", data$Total_Earnings))  
data$Number_of_Players <- as.numeric(gsub("[^0-9]", "", data$Number_of_Players))

data <- data %>%
  mutate(average_earning_per_player = Total_Earnings / Number_of_Players)

head(data)
write.csv(data, "country_overall.csv", row.names = FALSE)

```


Access the top 500 e-sports’ player earning from specific game in the past 365 days

```{r, warning = FALSE, message = FALSE}

urls <- c(
  "https://www.esportsearnings.com/players/highest-earnings-last-365-days",
  "https://www.esportsearnings.com/players/highest-earnings-last-365-days-top-200",
  "https://www.esportsearnings.com/players/highest-earnings-last-365-days-top-300",
  "https://www.esportsearnings.com/players/highest-earnings-last-365-days-top-400",
  "https://www.esportsearnings.com/players/highest-earnings-last-365-days-top-500"
)

fetch_tables <- function(url) {
  webpage <- read_html(url)
  tables <- webpage %>%
    html_nodes("table.detail_list_table") %>%
    lapply(html_table, fill = TRUE) %>%
    bind_rows()  
  
  countries <- webpage %>%
    html_nodes("table.detail_list_table img") %>%  
    html_attr("title")                             
  
  if (length(countries) == nrow(tables)) {
    tables$Country <- countries  
  } else {
    warning("Country count does not match table rows for URL: ", url)
    tables$Country <- NA  
  }
  return(tables)
}

all_data <- lapply(urls, fetch_tables)
final_table <- bind_rows(all_data)
head(final_table)
write.csv(final_table, "top500_365.csv", row.names = FALSE)

```

Additionally, we found the data in the horizontal time dimension (from 2018 to 2024),

```{r}

all_data <- data.frame(
  Year = integer(),
  Rank = character(),
  Country = character(),
  Total_Earnings = character(),
  Number_of_Players = character(),
  stringsAsFactors = FALSE
)

# from 2018 to 2024
for (year in 2018:2024) {
  
  url <- paste0("https://www.esportsearnings.com/history/", year, "/countries")
  webpage <- read_html(url)
  rows <- webpage %>% html_nodes("tr")
  data <- data.frame(
    Rank = character(),
    Country = character(),
    Total_Earnings = character(),
    Number_of_Players = character(),
    stringsAsFactors = FALSE
  )
  
  for (row in rows) {
    # grab data
    columns <- row %>% html_nodes("td") %>% html_text(trim = TRUE)
    if (length(columns) == 4) {
      data <- bind_rows(data, as.data.frame(t(columns), stringsAsFactors = FALSE))
    }
  }
  data$Year <- year
  all_data <- bind_rows(all_data, data)
  
}

colnames(all_data) <- c("Rank", "Country", "Total_Earnings", "Number_of_Players", "Year")
all_data$Year <- all_data$Rank
all_data[, 1:4] <- all_data[, 6:9]

all_data <- all_data[, c(5, 1:4)]
colnames(all_data) <- c("Year", "Current_year_Ranking", "Countries", "Overall_Earnings", "Number_of_Players")
all_data$Current_year_Ranking <- as.numeric(gsub("\\.", "", all_data$Current_year_Ranking))

head(all_data, 10)

write.csv(all_data, "Earnings_country_2018_2024.csv", row.names = FALSE)

```

Also, we access data about top games from 2018 to 2024,

```{r}

all_data <- data.frame(
  Year = integer(),
  Rank = integer(),
  Game = character(),
  Total_Earnings = character(),
  Total_Players = character(),
  Total_Tournaments = character(),
  stringsAsFactors = FALSE
)

# 2018 to 2024
for (year in 2018:2024) {
  url <- paste0("https://www.esportsearnings.com/history/", year, "/games")
  webpage <- read_html(url)
  rows <- webpage %>% html_nodes("tr")
  year_data <- data.frame(
    Rank = integer(),
    Game = character(),
    Total_Earnings = character(),
    Total_Players = character(),
    Total_Tournaments = character(),
    stringsAsFactors = FALSE
  )
  for (row in rows) {
    columns <- row %>% html_nodes("td") %>% html_text(trim = TRUE)
    if (length(columns) == 5) {
      year_data <- bind_rows(year_data, as.data.frame(t(columns), stringsAsFactors = FALSE))
    }
  }
  year_data$Year <- year
  all_data <- bind_rows(all_data, year_data)
}

all_data[, 6] <- all_data[, 1]
all_data[, 1:5] <- all_data[, 7:11]
all_data <- all_data[, 1:6]

colnames(all_data) <- c("Current_year_Rank", "Game", "Total_Earnings", "Total_Players", "Total_Tournaments", "Year")
all_data$Current_year_Rank <- as.numeric(gsub("\\.", "", all_data$Current_year_Rank))
head(all_data)
write.csv(all_data, "Earnings_topgame_2018_2024.csv", row.names = FALSE)

```

## Exploratory Analysis

```{r}

# Data import

country_overall <- read.csv("country_overall.csv")
esports_top100 <- read.csv("esports_earnings_top100_overall.csv")
top500_365 <- read.csv("top500_365.csv")
earnings_country <- read.csv("Earnings_country_2018_2024.csv")
earnings_topgame <- read.csv("Earnings_topgame_2018_2024.csv")

country_overall$Total_Earnings <- as.numeric(gsub("[\\$,]", "", country_overall$Total_Earnings))
esports_top100$Total_Earnings <- as.numeric(gsub("[\\$,]", "", esports_top100$Total_Earnings))
top500_365$`Total..Last.365.Days.` <- as.numeric(gsub("[\\$,]", "", top500_365$`Total..Last.365.Days.`))
earnings_country$Overall_Earnings <- as.numeric(gsub("[\\$,]", "", earnings_country$Overall_Earnings))
earnings_topgame$Total_Earnings <- as.numeric(gsub("[\\$,]", "", earnings_topgame$Total_Earnings))

```

