---
title: "Statistical Analysis"

output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, warning = FALSE, message = FALSE}

library(rvest)
library(dplyr)
library(tidyverse)
library(httr)
library(ggplot2)
library(tidyr)
library(maps)
library(knitr)
library(viridis)
library(broom)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 8,      
  fig.asp = 0.6,     
  out.width = "80%"   
)


theme_set(theme_minimal() + theme(legend.position = "right"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

rm(list = ls())

country_overall <- read.csv("data/country_overall.csv")
esports_top100 <- read.csv("data/esports_earnings_top100_overall.csv")
top500_365 <- read.csv("data/top500_365.csv")
earnings_country <- read.csv("data/Earnings_country_2018_2024.csv")
earnings_topgame <- read.csv("data/Earnings_topgame_2018_2024.csv")

country_overall$Total_Earnings <- as.numeric(gsub("[\\$,]", "", country_overall$Total_Earnings))
esports_top100$Total_Earnings <- as.numeric(gsub("[\\$,]", "", esports_top100$Total_Earnings))
top500_365$`Total..Last.365.Days.` <- as.numeric(gsub("[\\$,]", "", top500_365$`Total..Last.365.Days.`))
earnings_country$Overall_Earnings <- as.numeric(gsub("[\\$,]", "", earnings_country$Overall_Earnings))
earnings_topgame$Total_Earnings <- as.numeric(gsub("[\\$,]", "", earnings_topgame$Total_Earnings))
```

In this section, we conduct statistical modeling and analysis to investigate factors influencing total earnings in esports games. The analysis begins with t-tests to examine differences in earnings across countries over multiple years, assessing whether the industry demonstrated resilience during the COVID-19 pandemic by rebounding after a significant decline in 2020. This is followed by linear regression with interaction terms, residual diagnostics, and prediction visualization to further explore the factors impacting total earnings.

## 1.1 Data Cleaning and Feature Engineering for t-test

We consolidate earnings data across four years (2018, 2019, 2020, and 2021) by performing an iterative inner join on the Countries column. Each year’s data is first filtered, renamed for clarity (Earnings_<Year>), and included in the merge process. The resulting dataset aligns earnings for the same countries across all years, ensuring only countries present in all four years are included. This cleaned and structured data enables paired comparisons and year-over-year analysis.
```{r}
paired_data <- reduce(
  list(
    earnings_country %>% filter(Year == 2018) %>% select(Countries, Earnings_2018 = Overall_Earnings),
    earnings_country %>% filter(Year == 2019) %>% select(Countries, Earnings_2019 = Overall_Earnings),
    earnings_country %>% filter(Year == 2020) %>% select(Countries, Earnings_2020 = Overall_Earnings),
    earnings_country %>% filter(Year == 2021) %>% select(Countries, Earnings_2021 = Overall_Earnings)
  ),
  inner_join,
  by = "Countries"
)

```

## 1.2 Paired T-Tests
Null Hypothesis (H₀): There is no significant difference in esports earnings between 2018 (pre-pandemic) and 2019 (pandemic onset).
Alternative Hypothesis (H₁): Esports earnings in 2018 are significantly different from 2019, reflecting the impact of the pandemic.

Null Hypothesis (H₀): There is no significant difference in esports earnings between 2019 (pandemic onset) and 2020 (initial pandemic impact).
Alternative Hypothesis (H₁): Esports earnings in 2019 are significantly different from 2020, reflecting the industry's resilience or adaptation during the initial pandemic year.

Null Hypothesis (H₀): There is no significant difference in esports earnings between 2020 (pandemic onset) and 2021 (recovery phase).
Alternative Hypothesis (H₁): Esports earnings in 2021 are significantly different from 2020, indicating a recovery or continued decline.

```{r}
# Function for t-tests
perform_t_test <- function(data, col1, col2) {
  t_test <- t.test(data[[col1]], data[[col2]], paired = TRUE)
  tidy_results <- broom::tidy(t_test)
  return(tidy_results)
}

# Apply t-tests to year pairs
t_test_results <- list(
  `2018 vs 2019` = perform_t_test(paired_data, "Earnings_2018", "Earnings_2019"),
  `2019 vs 2020` = perform_t_test(paired_data, "Earnings_2019", "Earnings_2020"),
  `2020 vs 2021` = perform_t_test(paired_data, "Earnings_2020", "Earnings_2021")
)

# Combine results into a table
combined_results <- purrr::map_dfr(t_test_results, ~., .id = "Year Comparison")
combined_results %>%
  kable(digits = 5, caption = "Paired T-Test Results for Earnings Between Years")

```
### **Paired T-Test Results: Year-over-Year Changes in Esports Earnings (2018–2021)**

The paired t-test results reveal significant year-over-year changes in esports earnings between 2018 and 2021. Below are the key findings from the analysis:

#### **2018 to 2019: Significant Increase**
From 2018 to 2019, earnings increased significantly, with 2019 showing an average increase of **$545,207.8** compared to 2018 (*t = -3.37, p = 0.00104*). This growth indicates a steady expansion of the esports industry during this period, as the industry continued to attract more sponsorships, audiences, and competitive opportunities globally.

#### **2019 to 2020: Decrease During COVID-19 Pandemic**
During the initial year of the COVID-19 pandemic (2019 to 2020), earnings decreased by an average of **$772,820.2** (*t = 3.55, p = 0.00057*). Despite the increased popularity of online entertainment during lockdowns, the decline may be attributed to the disruption of large in-person events and delays in sponsorship commitments, which are crucial to the industry's revenue model.

#### **2020 to 2021: Post-Pandemic Recovery**
From 2020 to 2021, earnings increased significantly by an average of **$984,980.7** (*t = -3.18, p = 0.00192*). This rebound reflects the industry’s adaptation to post-pandemic conditions, with many events transitioning successfully to hybrid or online formats. However, the sharp increase also suggests potential challenges in sustaining growth as market dynamics normalized and traditional sports began to reclaim audience attention.


```{r}
# Pivot the paired_data to a long format
long_data <- paired_data %>%
  pivot_longer(
    cols = starts_with("Earnings_"),
    names_to = "Year",
    values_to = "Earnings"
  ) %>%
  mutate(
    Year = case_when(
      Year == "Earnings_2018" ~ "2018",
      Year == "Earnings_2019" ~ "2019",
      Year == "Earnings_2020" ~ "2020",
      Year == "Earnings_2021" ~ "2021"
    )
  )

# Create an interactive boxplot using plotly
plot_ly(
  data = long_data,
  x = ~Year,
  y = ~Earnings,
  type = "box",
  boxpoints = "all", # Include all points (jittered)
  jitter = 0.3,
  pointpos = -1.8,
  marker = list(color = "skyblue", opacity = 0.6)
) %>%
  layout(
    title = "Distribution of Esports Earnings by Year",
    xaxis = list(title = "Year"),
    yaxis = list(
      title = "Earnings",
      tickvals = seq(0, 100000, by = 20000), # Set y-axis ticks
      tickformat = ",.0f" # Format ticks as integers
    )
  )
```

The boxplot and paired t-test results reveal key insights into the year-over-year changes in esports earnings between 2018 and 2021. Below are the main observations and interpretations:

#### **1. Earnings Distribution: Not Symmetric**
- The boxplot indicates that earnings distributions for all years are **not symmetric**.
- Extreme outliers above the boxplots highlight that a few countries or organizations have disproportionately higher earnings compared to the majority.
- This asymmetry suggests a **skewed distribution**, where most countries earn relatively low amounts, while a small number achieve exceptionally high earnings.

#### **2. Mean vs. Median**
- For all years, the **mean** is likely higher than the **median**, as shown by the longer whiskers and outliers above the box.
- The paired t-tests measure the differences in **mean earnings**, which are heavily influenced by these extreme outliers, even though the median earnings may not change as significantly between years.


## 2.1 Data Cleaning and Feature Engineering for regression

We clean the data by removing special characters, converting variables to numeric types, and creating additional features such as game frequency and encoded year.Frequency Represents the number of times each game appears in the dataset across years.

```{r}

# Clean and process `Game` data
data <- earnings_topgame %>%
  mutate(
    Total_Earnings = as.numeric(gsub("[^0-9.]", "", Total_Earnings)),
    Total_Players = as.numeric(gsub("[^0-9]", "", Total_Players)),
    Total_Tournaments = as.numeric(gsub("[^0-9]", "", Total_Tournaments))
  ) %>%
  drop_na()  # Remove rows with missing values

# Count occurrences of each game
game_counts <- data %>%
  count(Game, name = "Frequency")

# Ensure both objects are data frames
data <- as.data.frame(data)
game_counts <- as.data.frame(game_counts)

# Perform a safe join
data <- left_join(data, game_counts, by = "Game")

# Add a numerical encoding for Year
data <- data %>%
  mutate(Year = as.numeric(as.factor(Year)))


```

## 2.2 Model Fitting with Interaction Terms

We conducted a linear regression analysis to model `log(Total_Earnings)` based on the following predictors: 
`Year`, `Total_Players`, `Total_Tournaments`, `Frequency`, and an interaction term (`Total_Players:Total_Tournaments`). Below is a summary of the key findings.

```{r}

data <- data %>%
  mutate(Total_Earnings = ifelse(Total_Earnings <= 0, 1, Total_Earnings))

data |> 
  lm(log(Total_Earnings) ~ Year + Total_Players * Total_Tournaments + Frequency, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 5)

```


```{r}
model_interaction <- lm(log(Total_Earnings) ~ Year + Total_Players * Total_Tournaments + Frequency, data = data)
summary(model_interaction)
```
#### Model interpretation:

All predictors, including the interaction term, are statistically significant (p<0.05).
Total_Players, Total_Tournaments, and Frequency show particularly strong effects.
The adjusted R-square value is approximately 0.42, indicating that the model explains 42% of the variation in log(Total_Earnings).

## 2.3 Model diagnosis

```{r}

# Residual plots
par(mfrow = c(1, 2))  # Arrange plots in a 1x2 grid

# Residuals vs Fitted
plot(model_interaction, which = 1, main = "Residuals vs Fitted")

# Normal Q-Q plot
plot(model_interaction, which = 2, main = "Normal Q-Q Plot")


```

The model provides significant insights into the predictors of esports earnings, with key variables such as Total_Players, Total_Tournaments, and Frequency showing strong effects. However, diagnostic plots reveal potential issues:

Non-linearity and Heteroscedasticity: Residuals suggest the model may not fully capture the relationship between predictors and earnings, especially for extreme values.
Outliers: Points like 219, 299, and 824 significantly deviate, indicating the need for robust methods or further investigation.
Normality: While most residuals align with normality, tail deviations highlight areas for improvement.

To refine the model, feature engineering, non-linear terms, and robust methods could enhance predictive performance and better handle anomalies.
